### 2022-02-04 plos_latex_template.tex
Submitted version

%Janani: for inserting sparingly few super relevant emojis
\usepackage{emoji} %Rocio: https://www.overleaf.com/learn/latex/Questions/Inserting_emojis_in_LaTeX_documents_on_Overleaf#Emoji_characters_and_the_Overleaf_editor

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*\Yinyang}, 
Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4,5,6},
Heather Turner\textsuperscript{7}, %\ddag},
Dorothea Hug Peter\textsuperscript{8}, %\ddag},
Natalia Soledad Morandeira\textsuperscript{9},
Matt Bannert\textsuperscript{10},
Batool Almazrouq\textsuperscript{11,12},
Liz Hare\textsuperscript{13},
Laura Ación\textsuperscript{5,14},
Juan Pablo Narváez-Gómez\textsuperscript{15},
Marcela Alfaro Córdoba\textsuperscript{16},
Federico Marini\textsuperscript{17},
Rita Giordano\textsuperscript{18},
Adithi Upadhya\textsuperscript{19},
Joselyn Chávez\textsuperscript{20},
Anicet Ebou\textsuperscript{21},
Silvia Canelón\textsuperscript{22},
Janani Ravi\textsuperscript{23}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Instituto de Pesquisas Jardim Botânico do Rio de Janeiro, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{3} International Institute for Sustainability, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina
\\
\textbf{5} MetaDocencia, Argentina
\\
\textbf{6} Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
\\
\textbf{7} Department of Statistics, University of Warwick, Coventry, UK
\\
\textbf{8} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
\\
\textbf{9} Instituto de Investigación e Ingeniería Ambiental, Universidad Nacional de San Martín - Consejo Nacional de Investigaciones Científicas y Técnicas, General San Martín, Buenos Aires, Argentina
\\
\textbf{10} KOF Swiss Economic Institute, ETH Zurich, Zurich, Switzerland
\\
\textbf{11} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK
\\
\textbf{12} King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
\\
\textbf{13} Dog Genetics LLC, Astoria, NY, USA
\\
\textbf{14} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
\\
\textbf{15} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brasil
\\
\textbf{16} Department of Statistics, University of California, Santa Cruz, CA, USA
\\
\textbf{17} Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center Mainz, Mainz, Germany
\\
\textbf{18} Royal Society of Chemistry, Thomas Graham House (290), Science Park, Milton Road, Cambridge, UK
\\
\textbf{19} ILK Labs/Bengaluru, Karnataka 560046, India
\\
\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
\\
\textbf{21} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
\\
\textbf{22} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
\\
\textbf{23} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
\\
\bigskip

@@

Conferences are spaces to meet and network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference. 
The authors recently organized the 2021 edition of the useR! statistical computing conference, attracting a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been mainly written for potential organizers or selection committees of conferences and contain multiple practical tips so that they can help diverse types of events to become as accessible and inclusive as they can and should be.

@@

% Conferences are great, but only for some
Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
A good conference experience can make a difference in the professional development of the participants and create long-lasting collaborations and opportunities. 
However, opportunities for participating in conferences are not equal for all. 
Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income countries, prestigious institutions, native English speakers with no disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
The authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
We had the opportunity to organize a virtual and global statistical computing conference, useR! 2021 \cite{sanchez-tapia_user_2021-2}, for users and developers of the R programming language \cite{r_core_team_2021}.
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.

% Roadmap of the paper
The rules result from our lessons learned before, during, and after the organization of the useR! conference. 
They are organized in three sections: three foundational rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundational rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and other spaces). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to encourage and ensure full participation of non-native English speakers. 
\textbf{Rule 8} provides tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion.

We believe that these rules will greatly benefit organizers who intend to put inclusion at the core of their conference right from the conception and planning phases.
The rules will also help people who are part of meeting committees that oversee the site/location selection process, or that coordinate with the local organizers of conferences. 
We have designed them for wide adoption to help scientific and technical conferences to become as accessible and inclusive as they can and should be.

@@

\begin{figure}[!h]
\centering
%ast: use Adobe Acrobat to add missing accessibility metadata to your PDF file. ask how to do this this when submitting - some edits to the PDF may be needed to add alt-text to the final version https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#late
\pdftooltip{
\includegraphics[width=\textwidth]{figs/rules.png}}{
%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: foundational rules (rules 1-3), design rules (rules 4 to 9), and a continuity rule (rule 10). The overall diagram depicts an arrow from left to right which contains all the rules. From left to right, "the foundations" label is surrounded by blue bubbles which show Rule "1 Vision", Rule "2 Safe environment", and Rule "3 Organizing Team". Then, the label "Design Rules" is surrounded by six blue bubbles with Rules "4 Counteract bias", "5 Online", "6 Accessibility", "7 Language", "8 Communication",  "9 Financial". To the right, and forming the head of the arrow, the Tenth Rule is under the label "Continuity" and consists of a blue bubble "10 Assess and pass along".
}
\caption{Schematic diagram of the rules organized in three groups: the foundations (Rules 1, 2, and 3), conference design (Rules 4 to 9), and continuity (Rule 10).}
\label{fig:diagram}
\end{figure}

\subsection*{Rule 1: Set a vision for diversity and inclusion}
\label{rule_diversity}

Diversity encompasses multiple dimensions: age, disability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. 
Nonetheless, we live in a world with implicit hierarchies along these axes. 
Some statuses (e.g. being cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems—including conferences—are consciously and unconsciously built. 
People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia, and sometimes these forms of oppression happen simultaneously and in complex ways \cite{crenshawDemarginalizingIntersectionRace1989}.
As a result, some groups of people have been systematically excluded from or only partially included in academia and scientific and professional circles.
Recognizing this systematic exclusion is paramount to inclusion.
Assess how it translates to your particular field and scientific or professional community and use this assessment to set the vision and goals of diversity and inclusion for your conference.
Define measurable indicators for these goals (e.g. a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions) that will help you be accountable along the way.
The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.

\subsection*{Rule 2: Create a safe and welcoming environment}
\label{rule_inclusion}

% Inclusion for everyone
% Explain why we need this
Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part of the conference activities and community. Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees. This means respecting all the aspects you committed to in your diversity vision (see \textbf{Rule 1}). For example, promote the use and respect of pronouns and commit to respect all genders and sexual orientations. If the conference has an in-person component, pre-assign quiet, private spaces for religious practices, lactation, and child care. Include menu options that account for diverse dietary and religious requirements.
Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and  great advice on `low-hanging fruits').
Open the schedule for events organized by thematic groups and communities (e.g. task force meetings, LGBTQIA+-friendly spaces, Black in STEM), and sessions aimed to welcome first-timers to the conference.

% A culture of support (active bystandership) and code of conduct for a safe place 
Creating a safe and welcoming environment also requires promoting a culture of support, including the idea that anyone can intervene respectfully in defense of other attendees in case anything unpleasant, incorrect, or offensive happens.
This is called `active bystandership'; consider providing training or pre-conference discussions to promote it.
However, the most important step that you should take to ensure the safety of the community is
adopting a code of conduct (CoC) and preparing a diverse response team to enforce it \cite{favaroYourScienceConference2016}.
The CoC is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behaviors, and the way to report violations \cite{auroraHowRespondCode2019}.
The CoC response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community. The team should be prepared early during conference planning because CoC violations can happen before the conference too (e.g. social media, organization meetings).  
We strongly recommend reading `How to Respond to code of conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC response team.

\subsection*{Rule 3: Have an inclusive and diverse organizing team}
\label{rule_organizing_team}

% Build a diverse team, and representation
A genuinely inclusive conference can only be organized by an inclusive and diverse team.
Organizers should comprise a team with people from as many different backgrounds as possible regarding regions, genders, ethnicities, career stages, and other aspects of diversity.
People with disabilities often say: `Nothing about us without us' \cite{charlton_nothing_1998, werner_nothing_1998}; the same holds for other dimensions of diversity. 
Having people from diverse backgrounds in decision-making roles will positively affect the conference as a whole because all the processes will benefit from their expertise, experience, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
The experience of people in marginalized groups
is especially important; they cannot be replaced by good intentions or second-hand knowledge from people who have not lived through the same experiences \cite{costanzachockDesign2020}.
In addition, a diverse team plays an essential role in creating a welcoming space because representation—seeing people with similar life experiences occupy critical positions of power and breaking negative stereotypes—is one of the best ways to create a sense of belonging for everyone participating in the conference.
It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work on diversity and inclusion aspects; 
every member should have the freedom to choose which areas of the conference they want to contribute to.

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
The regional and local communities, groups, or associations in your field are good sources to tap into. 
When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. 
People from underrepresented groups sometimes refrain to nominate themselves while people in dominant groups tend to do it more easily. 
Instead, nominate folks directly and offer them leading positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider paying them as an item in your budget (see \textbf{Rule 9}).
Furthermore, tasks such as receiving and responding to CoC reports can be emotionally intense work and should be additionally rewarded.
And most importantly, take care of your team and their well-being! Check on them regularly and make sure that everyone is comfortable. This represents a great amount of work and you will need to support each other in the long run.

\subsection*{Rule 4: Consciously counteract bias in the conference program}
\label{rule_unbias}

The conference program may lack diversity in the invited roles––such as keynote speakers, program/scientific committee, session chairs, panelists––, the accepted abstracts, and even the scope of the topics covered. 
It is important to counteract biases in all of these aspects.
When choosing or inviting people for visible and valued roles in the conference it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. 
Rather than deter you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.

Biases in the final list of abstracts begin with self-selection: some potential conference participants may not feel confident enough to submit their work, especially to large and prestigious conferences.
To mitigate this, 
some communities have pre-submission mechanisms to aid their members prepare and receive feedback for their abstract (e.g. the Latin-R and R-Ladies communities have volunteers to help their members improve their conference abstracts).
If this does not exist in your field yet, consider asking members of your community to provide a similar kind of mentorship.

The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted. 
Selection committees need to be reminded of such biases in the evaluation process to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups, or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019}.
Evaluate the plausibility of anonymizing all of the submission information to
define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
Define clear evaluation criteria and share them with the reviewers and authors.
Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe; even well-meaning and carefully crafted processes can be prone to reproduce the biases observed in academia and society.

@@

\subsection*{Rule 5: Design a strong online component} 
\label{rule_online}

>>>>

%Janani: for inserting sparingly few super relevant emojis
% \usepackage{emoji} %Rocio: https://www.overleaf.com/learn/latex/Questions/Inserting_emojis_in_LaTeX_documents_on_Overleaf#Emoji_characters_and_the_Overleaf_editor

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*\Yinyang}, 
Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4,5,6},
Heather Turner\textsuperscript{7}, %\ddag},
Dorothea Hug Peter\textsuperscript{8}, %\ddag},
Natalia Soledad Morandeira\textsuperscript{9},
Matt Bannert\textsuperscript{10},
Batool Almazrouq\textsuperscript{11,12},
Liz Hare\textsuperscript{13},
Laura Ación\textsuperscript{5,14},
Juan Pablo Narváez-Gómez\textsuperscript{15},
Marcela Alfaro Córdoba\textsuperscript{16},
Federico Marini\textsuperscript{17},
Rita Giordano\textsuperscript{18},
Silvia Canelón\textsuperscript{19},
Anicet Ebou\textsuperscript{20},
Adithi Upadhya\textsuperscript{21},
Joselyn Chávez\textsuperscript{22},
Janani Ravi\textsuperscript{23}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Instituto de Pesquisas Jardim Botânico do Rio de Janeiro, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{3} International Institute for Sustainability, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina
\\
\textbf{5} MetaDocencia, Argentina
\\
\textbf{6} Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
\\
\textbf{7} Department of Statistics, University of Warwick, Coventry, UK
\\
\textbf{8} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
\\
\textbf{9} Instituto de Investigación e Ingeniería Ambiental, Universidad Nacional de San Martín - Consejo Nacional de Investigaciones Científicas y Técnicas, General San Martín, Buenos Aires, Argentina
\\
\textbf{10} KOF Swiss Economic Institute, ETH Zurich, Zurich, Switzerland
\\
\textbf{11} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK
\\
\textbf{12} King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
\\
\textbf{13} Dog Genetics LLC, Astoria, NY, USA
\\
\textbf{14} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
\\
\textbf{15} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brasil
\\
\textbf{16} Department of Statistics, University of California, Santa Cruz, CA, USA
\\
\textbf{17} Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center Mainz, Mainz, Germany
\\
\textbf{18} Royal Society of Chemistry, Thomas Graham House (290), Science Park, Milton Road, Cambridge, UK
\\
\textbf{19} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
\\
\textbf{20} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
\\
\textbf{21} ILK Labs/Bengaluru, Karnataka 560046, India
\\
\textbf{22} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
\\
\textbf{23} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
\\
\bigskip

@@

Conferences are spaces to meet and network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference based on the authors' recent experience organizing the 2021 edition of the useR! statistical computing conference, which attracted a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been mainly written for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive. We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.

@@

% Conferences are great, but only for some
Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share recent contributions.
A good conference experience can make a difference in the professional development of the participants and create long-lasting collaborations and opportunities. 
However, opportunities for participating in conferences are not equally available for all. 
Many academic and tech conferences have been spaces that reproduce systemic inequalities, by failing to overcome the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income backgrounds, prestigious institutions, native English speakers with no disabilities) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
Like many others, the authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
We had the opportunity to organize a virtual and global statistical computing conference, useR! 2021 \cite{sanchez-tapia_user_2021-2}, for users and developers of the R programming language \cite{r_core_team_2021}.
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.

% Roadmap of the paper
Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference. 
They are organized in three sections: three foundational rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundational rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and other spaces). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to encourage and ensure full participation of non-native English speakers. 
\textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion, and passing the torch to future organizers.

We believe that these rules can guide organizers who intend to put inclusion at the core of their conference right from the conception and planning phases.
The rules will also help people who are part of meeting committees that oversee the site/location selection process, or that coordinate with the local organizers of conferences. 
We see the rules as a starting point for conversations and efforts towards building more inclusive conferences across the world.

@@

\begin{figure}[!h]
\centering
%ast: use Adobe Acrobat to add missing accessibility metadata to your PDF file. ask how to do this this when submitting - some edits to the PDF may be needed to add alt-text to the final version https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#late
\pdftooltip{
\includegraphics[width=\textwidth]{figs/fig1_rules.png}}{
%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: foundation rules (Rules 1-3), design rules (Rules 4 to 9), and a continuity rule (Rule 10). The overall diagram depicts an arrow from left to right which contains all the rules. In the left column, under the first label" Foundation", there are three text boxes in light blue: "1. Vision", "2. Safe environment", and "3. Organizing Team". In the middle column, under the label "Design Rules" there are six light green boxes with Rules "4. Counter bias", "5. Virtual / hybrid", "6. Accessibility", "7. Language", "8. Communication",  "9. Finance". To the right, below the arrowhead labeled Continuity, is the Tenth Rule in a light pink box "10. Pass the torch".
}
\caption{Schematic diagram of the rules organized in three groups: foundation (Rules 1, 2, and 3), design (Rules 4 to 9), and continuity (Rule 10).}
\label{fig:diagram}
\end{figure}

\section*{Rule 1: Set a vision for diversity and inclusion}
\label{rule_diversity}

Diversity encompasses multiple dimensions: age, disability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. 
Nonetheless, we live in a world with implicit hierarchies along these axes. 
Some statuses (e.g. being cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems—including conferences—are consciously and unconsciously built. 
People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia, and sometimes these forms of oppression happen simultaneously and in complex ways (a concept known as `intersectionality' \cite{crenshawDemarginalizingIntersectionRace1989}).
As a result, some groups of people have been systematically excluded from or only partially included in academic, scientific, and professional circles.
Recognizing this systematic exclusion is paramount to inclusion.
Assess how lack of diversity translates to your particular field and scientific or professional community and use this assessment to set the vision and goals of diversity and inclusion for your conference.
Define measurable indicators for these goals (e.g. a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions) that will help you be accountable along the way.
The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.

\section*{Rule 2: Create a safe and welcoming environment}
\label{rule_inclusion}

% Inclusion for everyone
% Explain why we need this
Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part of the conference activities and community. Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees. This means respecting all the aspects included in your diversity vision (see \textbf{Rule 1}). For example, promote the use and respect of pronouns and commit to respect all genders and sexual orientations. If the conference has an in-person component, pre-assign quiet, private spaces for religious practices, lactation, and child care. Include menu options that account for diverse dietary and religious requirements.
Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and great advice on `low-hanging fruits'), some others need careful planning in advance, so it's good to have them in mind from the beginning.
Open the schedule for events organized by thematic groups and communities (e.g. task force meetings, LGBTQIA+-friendly spaces, Black in STEM), and sessions aimed to welcome first-timers to the conference.

% A culture of support (active bystandership) and code of conduct for a safe place 
Creating a safe and welcoming environment also requires promoting a culture of support, including the idea that anyone can intervene respectfully in defense of other attendees in case anything unpleasant, incorrect, or offensive happens.
This is called `active bystandership'; consider providing training or pre-conference discussions to promote it.
However, the most important step that you should take to ensure the safety of the community is
adopting a code of conduct (CoC) and preparing a diverse response team to enforce it \cite{favaroYourScienceConference2016}.
The CoC is a living document meant to keep the community safe and should state clearly the unacceptable behaviors, the consequences for engaging in such behaviors, and the way to report violations \cite{auroraHowRespondCode2019}.
The CoC response team must receive training on how to receive reports, respond to incidents, and communicate their responses to the whole community. The team should be prepared early during conference planning because CoC violations can happen before the conference too (e.g. social media, organization meetings).  
Responding to CoC reports can be one of the most emotionally intense tasks of the team. Resources permitting, consider adding their work to your budget (see \textbf{Rule 9}).
We strongly recommend reading `How to Respond to code of conduct Reports' \cite{auroraHowRespondCode2019} as an excellent starting point for building a strong CoC response team.

\section*{Rule 3: Gather an inclusive and diverse organizing team}
\label{rule_organizing_team}

% Build a diverse team, and representation
A genuinely inclusive conference can only be organized by an inclusive and diverse team.
Organizers should comprise a team with people from as many different backgrounds as possible in terms of regions, genders, ethnicities, career stages, and other aspects of diversity.
People with disabilities often say: `Nothing about us without us' \cite{charlton_nothing_1998, werner_nothing_1998}; the same holds for other dimensions of diversity. 
Having people from diverse backgrounds in decision-making roles will positively affect the conference as a whole because all the processes will benefit from their expertise, experience, and distinct perspectives \cite{hongGroupsDiverseProblem2004}. 
The experience of people in marginalized groups
is especially important; they cannot be replaced by good intentions or second-hand knowledge from people who have not lived through the same experiences \cite{costanzachockDesign2020}.
In addition, a diverse team plays an essential role in creating a welcoming space because representation—seeing people with similar life experiences occupy critical positions of power and breaking negative stereotypes—is one of the best ways to create a sense of belonging for everyone participating in the conference.
It is important, though, to ensure that organizers with diverse backgrounds are not restricted to only work on diversity and inclusion aspects; 
every member should have the freedom to choose which areas of the conference they want to contribute to.

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
The regional and local communities, groups, or associations in your field are good sources to tap into. 
When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably.
Instead, nominate folks directly and offer them leadership positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider reallocating resources to give them a stipend for their work (see \textbf{Rule 9}).
In any case, try to find meaningful ways to acknowledge everybody's work. Use the conference spaces (e.g. webpage, opening/closing ceremony at the conference, social media) to give proper credit to the people who are making the event possible.
And most importantly, take care of your team and their well-being! Check on them regularly and make sure that everyone is comfortable. This represents a great amount of work and you will need to support each other in the long run.

\section*{Rule 4: Consciously counteract bias in the conference program}
\label{rule_unbias}

The conference program may lack diversity in the invited roles––such as keynote speakers, program/scientific committee, session chairs, panelists––, the accepted abstracts, and even the scope of the topics covered. 
It is important to counteract biases in all of these aspects.
When choosing or inviting people for visible and valued roles in the conference, it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. 
Rather than deterring you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.

Biases in the final list of abstracts begin with self-selection: some potential conference participants may not feel confident enough to submit their work, especially to large and prestigious conferences.
To mitigate such self-selection, some communities have pre-submission mechanisms to aid their members prepare and receive feedback for their abstract (e.g. the Latin-R and R-Ladies communities have volunteers to help their members improve their conference abstracts).
If this does not exist in your field yet, consider asking members of your community to provide a similar kind of mentorship.

The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted. 
Selection committees need to be reminded of such biases in the evaluation process to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups, or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
Evaluate the plausibility of anonymizing all of the submission information to
define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
Define clear evaluation criteria and share them with the reviewers and authors.
Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe; even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.

@@

\section*{Rule 5: Design a strong online component} 
\label{rule_online}
